---
title: "Biomass model - displaying species-level trends"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr, warn.conflicts = FALSE)
library(ggplot2)
library(readxl)
library(stringr)
library(tidyr)


# Use runjags because it is a little easier to run parallel MCMC chains
# than using rjags directly.
library(runjags)

runjags.options(modules = "glm", method = "parallel")

# Default graph theme
theme_set( theme_bw() )


# Function to calculate highest posterior density intervals
# (adapted from coda::HPDinterval). Returns a matrix with a 
# row for each input variable.
hpdi <- function (x, prob = 0.95) 
{
  x <- as.matrix(x)
  Nr <- nrow(x)
  if (Nr <= 1) stop("x must have more than 1 vector element or matrix row")
  x <- apply(x, 2, sort)

  Nc <- ncol(x)
  gap <- max(1, min(Nr - 1, round(Nr * prob)))
  init <- 1:(Nr - gap)
  
  inds <- apply(x[init + gap, , drop = FALSE] - x[init, , drop = FALSE], 
                MARGIN = 2, 
                which.min)
  
  out <- cbind(x[cbind(inds, 1:Nc)], x[cbind(inds + gap, 1:Nc)])
  colnames(out) <- c("lower", "upper")
  
  rnms <- colnames(x)
  if (is.null(rnms)) rnms <- paste0("X", 1:Nc)
  rownames(out) <- rnms
  
  out
}

```


## Summary

This is a modification of the total biomass model (see document `TB2AUG20.Rmd`) to fit and display species-level effects within each cell (veg x watering x CO2 combination).

*Caution!* This is essentially just for graphical exploration of the previously fitted, veg-level model. It is not a fully hierarchical model because there is no information sharing between species and veg type levels.


## Data

```{r}

DAT <- read_excel("BIOMASS_DATA_2AUG20.xlsx") %>%
  mutate_if(is.character, as.factor) %>%
  rename(watering = drought) %>%
  
  # subset data to records without missing biomass values
  dplyr::filter(!is.na(bg_total)) %>%
  dplyr::filter(bg_total > 0) %>%
  
  dplyr::select(veg, co2, watering, species, bg_total) %>%
  
  # Assign cell indices and labels
  group_by(veg, watering, co2) %>%
  
  mutate(cell = cur_group_id(),
         cell_label = paste0(
           ifelse(co2 == "400ppm", "400", "640"),
           ifelse(watering == "drought", "Dr", "Wtr"),
           veg
         )) %>%
  
  ungroup() %>%

  # Assign index (1-3) for species within veg type. This 
  # can be used as species within treatment cell.
  group_by(veg) %>%
  mutate(sp_in_cell = as.integer(factor(species))) %>%
  ungroup() %>%
  
  # Assign extended labels to species to make graphs easier to read
  mutate(species_veg = sprintf("%s: %s", veg, species),
         species_veg = factor(species_veg))

head(DAT)

```


Quick graph of the data:

```{r}

ggplot(data = DAT) +
  geom_point(aes(x = bg_total, y = cell_label, colour = species_veg),
             position = position_dodge(width = 0.6),
             size = 2) +
  
  scale_colour_brewer(name = "Species", palette = "Set1") +
  
  labs(x = "Total biomass (g)", y = "")

```

## Model

### JAGS model code

Note that we provide two identical vectors for the response variable (total biomass) to the model, denoted as `y` and `ycopy`. This is to avoid JAGS complaining that a single response variable is being modelled in two, unconnected ways.

```{r}

modelTxt <- "model {
  # Cell means
  for (i in 1:length(y)) {
    # Responses modelled as Gamma distributed values
    y[i] ~ dgamma(shape[i], rate[i])
    
    # Relate parameters of Gamma distribution to
    # cell means and variances
    shape[i] <- mu[cell[i]]^2 / sd[cell[i]]^2
    rate[i] <- mu[cell[i]] / sd[cell[i]]^2
  }
  
  # Species-within-cell means
  # Note: there is no connection to the cell means (veg-level) model above.
  for (i in 1:length(ycopy)) {
    # Responses modelled as Gamma distributed values
    ycopy[i] ~ dgamma(shape.sp[i], rate.sp[i])
    
    # Relate parameters of Gamma distribution to
    # cell means and variances
    shape.sp[i] <- mu.sp[cell[i], sp_in_cell[i]]^2 / sd.sp[cell[i], sp_in_cell[i]]^2
    rate.sp[i] <- mu.sp[cell[i], sp_in_cell[i]] / sd.sp[cell[i], sp_in_cell[i]]^2
  }
  
  # Priors for cell means and standard deviations at the
  # veg type level
  for (k in 1:max(cell)) {
    # Prior means are constrained to be positive
    mu[k] ~ dnorm(0, 1e-3)I(0, )
    sd[k] ~ dunif(1, 30)
  }
  
  # Priors for cell means and standard deviations at the
  # species-within-veg type level
  for (k in 1:max(cell)) {
    for (isp in 1:3) {
      # Prior means are constrained to be positive
      mu.sp[k, isp] ~ dnorm(0, 1e-3)I(0, )
      sd.sp[k, isp] ~ dunif(1, 30)
    }
  }
  
}"

```


### Prepare data and run model

Add an integer index for species-within-cell.

```{r}

dat.model <- DAT %>%
  # Note: creating two copies of the response data
  dplyr::select(y = bg_total,
                ycopy = bg_total,
                cell, sp_in_cell) %>%
  
  arrange(cell, sp_in_cell) 
  

```


```{r}

model <- run.jags(modelTxt, 
                  data = dat.model, 
                  monitor = c("mu", "mu.sp"), 
                  n.chains = 4, 
                  adapt = 1000, 
                  burnin = 4000, 
                  sample = 2000,
                  thin = 10)

```

Quick check of model convergence by examining the number of effectively independent samples for each model parameter. Because there are not many parameters we can just use the run.jags::summary function for this.

```{r}

s <- summary(model)
summary( s[, "SSeff"] )

```

### Model results

Calculate the 95% highest posterior density interval for each of the `mu` (cell mean) and `mu.sp` (species-within-cell mean) parameters.

```{r}

# Combine model chains into a single matrix
post <- do.call(rbind, model$mcmc)

# 50% posterior intervals
post.50 <- hpdi(post, prob = 0.5) %>%
  as.data.frame() %>%
  mutate(param = rownames(.)) %>%
  rename(lower50 = lower, upper50 = upper)

rownames(post.50) <- NULL

# 95% posterior intervals
post.95 <- hpdi(post, prob = 0.95) %>%
  as.data.frame() %>%
  mutate(param = rownames(.)) %>%
  rename(lower95 = lower, upper95 = upper)

rownames(post.95) <- NULL

post.intervals <- left_join(post.50, post.95, by = "param") %>%
  dplyr::select(param, lower95, lower50, upper50, upper95)
  
```

```{r}
DAT <- DAT %>%
  tidyr::unite(watering_co2, watering, co2) 
```

```{r}
DAT1 <- DAT %>%
  mutate(species_veg = factor(species_veg, levels = c("DSF: C. gummifera", "DSF: E. pilularis", "DSF: E. sieberi", "GWD: E. blakelyi", "GWD: E. bridgesiana", "GWD: E. melliodora", "WSF: C. maculata", "WSF: E. fastigata", "WSF: E. viminalis"),
                              labels = c("DSF: C. gummifera", "         E. pilularis", "         E. sieberi", "GW: E. blakelyi", "         E. bridgesiana", "         E. melliodora", "WSF: C. maculata", "          E. fastigata", "          E. viminalis")))
```

Graph the intervals for cell and species-within-cell means against the observed data.

```{r fig.height = 2.25, fig.width = 4}

#xtext <- c(expression("D" ~ aCO^{2}), expression("D" ~ eCO^{2}), expression("W" ~ aCO^{2}), expression("W" ~ eCO^{2}))

cell_labels <- DAT1 %>%
  distinct(cell, cell_label)

cell_species_labels <- DAT1 %>%
  distinct(cell, cell_label, sp_in_cell, species_veg) %>%
  arrange(cell, sp_in_cell)

dat.gg.cell <- post.intervals %>%
  dplyr::filter(str_detect(param, "^mu\\[")) %>%
  mutate(cell = as.integer( str_extract(param, "\\d+") )) %>%
  left_join(cell_labels)

dat.gg.species <- post.intervals %>%
  dplyr::filter(str_detect(param, "^mu.sp")) %>%
  mutate(indices = str_extract(param, "\\d+,\\d+")) %>%
  tidyr::separate(indices, into = c("cell", "sp_in_cell"), convert = TRUE) %>%
  left_join(cell_species_labels)


My_Theme = theme(
  plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
  axis.title.y = element_text(size = 14, face = "bold", hjust = 0.5, vjust = 2.5),
  axis.text.y = element_text(size = 14, margin = unit(c(0.1, 0.1, 0.1, -0.2), "cm")),
  axis.text.x = element_text(size = 14, angle = 60, margin = unit(c(0.1, 0.1, 0.1, -0.2), "cm"), hjust = 1),
  axis.title.x = element_text(size = 14, face = "bold"),
  strip.text.x = element_text(size = 14),
  strip.text.y = element_text(size = 14),
  strip.text = element_text(face = "bold"),
  strip.background = element_blank(),
  panel.background = element_blank(),
  axis.ticks.length = unit(-0.07, "cm"),
  axis.ticks = element_line(size = 0.8),
  panel.grid.minor.x = element_blank(),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.y = element_blank(),
  legend.position = c(0.125, 0.75),
  legend.text.align = 0,
  legend.spacing.x =  unit(0.01, 'cm'),
  legend.text = element_text(size = 8, hjust = 0.5, face = 'italic'),
  plot.margin = unit(c(5, 5, 5, 5), "pt"))

ytext <- c(expression("DSF: D+" ~ aCO[2]), expression("GWD: D+" ~ aCO[2]), expression("WSF: D+" ~ aCO[2]), expression("DSF: W+" ~ aCO[2]), expression("GWD: W+" ~ aCO[2]), expression("WSF: W+" ~ aCO[2]), expression("DSF: D+" ~ eCO[2]), expression("GWD: D+" ~ eCO[2]), expression("WSF: D+" ~ eCO[2]), expression("DSF: W+" ~ eCO[2]), expression("GWD: W+" ~ eCO[2]), expression("WSF: W+" ~ eCO[2]))

colours <- c("springgreen4", "springgreen3", "seagreen2", "firebrick4", "orangered2", "darksalmon", "blue4", "royalblue3", "steelblue1")

dodgew <- 0.2

ggBGB <- ggplot(data = DAT1, aes(y = cell_label)) +
  geom_point(aes(x = bg_total, colour = species_veg),
             position = position_dodge(width = dodgew),
             size = 3, shape = 1, alpha = 0.6) +
  
  My_Theme +
  
  # Species level intervals
  geom_linerange(data = dat.gg.species,
                 aes(xmin = lower50, xmax = upper50, colour = species_veg),
                 position = position_dodge(width = dodgew),
                 size = 3, alpha = 0.75) +
  
  geom_linerange(data = dat.gg.species,
                 aes(xmin = lower95, xmax = upper95, colour = species_veg),
                 position = position_dodge(width = dodgew),
                 size = 1.5, alpha = 0.5) +
  
  # Cell level intervals
  geom_linerange(data = dat.gg.cell,
                 aes(xmin = lower50, xmax = upper50),
                 size = 3.5, alpha = 1) +
  
  geom_linerange(data = dat.gg.cell,
                 aes(xmin = lower95, xmax = upper95),
                 size = 1.5, alpha = 0.75) +
  
  scale_y_discrete(labels=ytext) +
  
  scale_colour_manual(values = colours) +
  
  labs(x = "Below-ground biomass (g)", y = "Treatment combination", colour = "Species") + 
  
  coord_flip()


ggsave(filename = "bg_total_cell_species.png")

print(ggBGB)

```


